---
title: "simpleRegress"
author: "Elisa Gonzalez"
date: "Fall 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## Model WEIGHT ~ AGE  this is an example 
```{r}
agemodel <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(simp)
summary.lm(simp)
```
1. Neck--> ARsqrd: 0.87  RST:49.91
2. Age-->  ARsqrd : 0.55 RST:81.47

The agemodel graphs plots are further away from te line than the neckmodel. Making the age model worst than that neckmodel. 


## Model WEIGHT ~ HEADLEN  this is an example 

```{r}
headlenmodel<- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

## Model WEIGHT ~ HEADWTH this is an example 

```{r}
headwthmodel <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

## Model WEIGHT ~ Length  this is an example 

```{r}
lengthmodel <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp)
summary.lm(simp)
```

```{r}
chestmodel <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

Best to Worst 
1. Chest        ARsqrd: 0.92.  RST:33.08
2. Neck         ARsqrd: 0.07    RST:49.91
3. Length       ARsqrd: 0.74    RST:61.83
4. Head length  ARsqrd: 0.69    RST:67.81
5. Head width   Arsqrd: 0.60    RST:76.41
6. Age          ARsrd: 0.55     RST:81.47  

After completing all the models, the best model is the chestmodel. The Chest adjusted R squared is higher than the other models, while the standard error is low. As seen in the graph, since most dots are closet to the line. As the lost continues on the plots of the models are further away from the lines than the chestmodels. Making the standard error higher and adjusted R squared lower. 